---
title: "Parameter Descriptions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parameter Descriptions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Parameters

ModVege makes use of a sizable number of different variables. Some of them 
are used as inputs to the simulation, some tune model behaviour and others 
are the actual state variables of the simulated grassland system. 

The following section attempt to provide an as comprehensive as possible list 
of all of these. For a thorough understanding, the reader is referred to the 
original publication by [@jouven2006ModelPredictingDynamics].

### State Variables

These values evolve during a model run and are stored in the respective 
fields in the [ModvegeSite] instance that was run.

```{r child = "children/state_variables.Rmd"}
```

### Initial conditions
```{r child = "children/initial_conditions.Rmd"}
```

### Model parameters
```{r child = "children/parameters.Rmd"}
```

### Functional group parameters
```{r child = "children/parameters_fg.Rmd"}
```

### Weather variables
```{r child = "children/weather_parameters.Rmd"}
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

